Verifying Spatial Properties of Stencil Computations

### 2016-11-15 paper.tex
shrunk section 4 considerably

  \emph{Absolute index} is a constant integer expression relative to the
  enclosing loop.
\end{definition}

@@

\begin{definition}[Stencil computations]
\label{def:stencil}
  Let $a$ be an array of dimensionality $n$ and $\overline{b}$ a
  collection of arrays of arbitrary (possibly differing)
  dimensionalities.
  A \emph{stencil computation} comprises an iteration by a set $I$ of at
  most $n$ induction variables
  over a subset of the index space of $a$. Each iteration
  determines the elements of $a$ by an
  assignment $a(\bar{e}) = e_r$ where each $e \in
  \bar{e}$ is either a neighbourhood or constant index.
  For all $b \in \bar{b}$ and array subscripts $b(\bar{e}')$
  that flow to $e_r$, each $e' \in \bar{e}'$ is
  either a neighbourhood or absolute index.

  We refer to the assignment $a(\bar{e}) = e_r$ and all the associated
  statements with dataflow to $e_r$ as the \emph{stencil kernel}.
\end{definition}
%
% specifying the shape of indexing expressions for
%arrays $\bar{v}$ that are simultaenously permitted and expected within
%the code to which the specification is given.
%
%\marginpar{\dnote{Is it worth having this informal consistency here?}}

Now we informally describe what it means for a stencil kernel to be consistent
with a specification. This will will be formalised in \S~\ref{sec:semantics} and
algorithmically described in  and \S~\ref{sec:analysis}.
Specifications are associated to a stencil kernel and are declared for one or
more array variables $\bar{v}$ within that kernel.  A stencil specification
defines a \emph{region} associated to an array variable $a$ against which a
stencil kernel should be \emph{consistent}. The region defined by a
specification is an $n$-dimensional rectilinear shape in a discrete space
crossing the origin with respect to the induction variable. Consistency is then
a two way requirement. All array subscripts on $a$
contributing to the kernel should be consistent with the region. Conversely,
all parts of the region should have an index corresponding to that part.

@@

Figure~\ref{fig:syntax} gives the syntax of stencil specifications, which is
detailed below. The entry point is the \nonterm{specification} production which
splits into either a \emph{region declaration} or a \emph{specification
declaration}.

@@

Specifications are subject to a set of equations which can be used to
understand when specifications are equivalent. Furthermore,
specifications are subject to a notion of approximation. We outline
the equational theory and notion of approximation here.

\begin{figure}
\vspace{-0.3em}
\begin{align*}
\setlength{\arraycolsep}{0.05em}
\begin{array}{c}
%\framebox{$\textit{region} \equiv \textit{region}'$} \\[0.6em]
\begin{array}{rll}
(\textsc{F\, \texttt{+} \,F}) \;\; &
\stenFwdSR{n \, \textsf{max} \, m}{d}{p \vee q} \\
 \equiv \; & \stenFwdSR{n}{d}{p} \; \texttt{+} \; \stenFwdSR{m}{d}{q} \\[0.6em]
%
%
(\textsc{B\, \texttt{+} \,B}) \;\; &
\stenBwdSR{n \, \textsf{max} \, m}{d}{p \vee q} \\
 \equiv \; & \stenBwdSR{n}{d}{p} \; \texttt{+} \; \stenBwdSR{m}{d}{q} \\[0.6em]
%
%
(\textsc{C\, \texttt{+} \,C}) \;\; &
\stenCenSR{n \, \textsf{max} \, m}{d}{p \vee q} \\
\equiv \; & \stenCenSR{n}{d}{p} \; \texttt{+} \; \stenCenSR{m}{d}{q} \\[0.6em]
%
%
(\textsc{C\, \texttt{+} \,F}) \;\; & \stenCenSR{n}{d}{p \vee q} \\
\textit{$m \leq n$} \; \equiv \; & \stenCenSR{n}{d}{p} \; \texttt{+} \;
                      \stenFwdSR{m}{d}{q} \\[0.6em]
%
%
(\textsc{C\, \texttt{+} \,B}) \;\; &
\stenCenSR{n}{d}{p \vee q} \\
\textit{$m \leq n$} \; \equiv \; & \stenCenSR{n}{d}{p} \; \texttt{+} \;
                      \stenBwdSR{m}{d}{q} \\[0.6em]
%
%
(\textsc{B\, \texttt{+} \,F}) \;\; &
\stenCenSR{n}{d}{p \vee q} \\
\equiv \; & \stenFwdSR{n}{d}{p} \; \texttt{+} \; \stenBwdSR{n}{d}{q}
  \\[0.6em]
%
%
(\textsc{P\, \texttt{+} \,F}) \;\; &
\stenFwdS{n}{d} \\
\equiv \; & \stenReflSR{d} \; \texttt{+} \; \stenFwdSR{n}{d}{p} \\[0.6em]
%
%
(\textsc{P\, \texttt{+} \,B}) \;\; &
\stenBwdS{n}{d} \\
\equiv \; & \stenReflSR{d} \; \texttt{+} \; \stenBwdSR{n}{d}{p} \\[0.6em]
%
%
(\textsc{P\, \texttt{+} \,C}) \;\; &
\stenCenS{n}{d} \\
\equiv \; & \stenReflSR{d} \; \texttt{+} \; \stenCenSR{n}{d}{p} \\[0.6em]
%
%
(\textsc{npoint}) \;\; & (R \, \term{*} \, S^\irreflS) \, \term{+} \,
                         (R^\irreflS \, \term{*} \, S) \;
\equiv \; R \, \term{*} \, S \\[0.4em]
%
(\textsc{$\ast$Dim}) \;\; & R^d \, \term{$\ast$} \, S^d \equiv R^d
                             \, \term{+} \, S^d \\[0.4em]
%
(\texttt{+}\textsc{idem}) \;\; & S \; \texttt{+} \; S \; \equiv \; S \\[0.4em]
%
(\texttt{+}\textsc{comm}) \;\; & S \; \texttt{+} \; T \; \equiv \; T \;
                       \texttt{+} \; S \\[0.4em]
%
(\texttt{+}\textsc{assoc}) \;\; & R \; \texttt{+} \; (S \; \texttt{+} \; T) \; \equiv \; (R \;
                       \texttt{+} \; S) \; \texttt{+} \; T \\[0.4em]
(\texttt{*}\textsc{comm}) \;\; & S \; \texttt{*} \; T \; \equiv \; T \;
                       \texttt{*} \; S \\[0.4em]
%
(\texttt{*}\textsc{assoc}) \;\; & R \; \texttt{*} \; (S \; \texttt{*} \; T) \; \equiv \; (R \;
                       \texttt{*} \; S) \; \texttt{*} \; T \\[0.4em]
(\textsc{dist}) \;\; & R \; \texttt{*} \; (S \; \texttt{+} \; T) \; \equiv \; (R \;
                       \texttt{*} \; S) \; \texttt{+} \; (R
                       \; \texttt{*} \; T)
\end{array}
\end{array}
\end{align*}
\caption{Equations on specifications (\emph{regions})}
\vspace{-0.5em}
\label{fig:equations}
\end{figure}

Figure~\ref{fig:equations} lists the equational theory for
specifications given by the binary relation $\equiv$
on \emph{region} terms. The equations use abbreviations
\term{point}, \term{fwd}, \term{bwd}, \term{cen}
for \term{nonpointed}, \term{forward}, \term{backward},
\term{centered} respectively. Furthermore, the abbreviations include syntactic sugar
for quantifying over the absence or presence of the \irrefl{}
attribute on region constants as a boolean. For example,
\term{fwd} syntactic sugar is defined:
%%
\begin{align*}
\hspace{-0.25em}
& \stenFwdSR{n}{d}{r}
\; := \; \\[-0.4em]
& \qquad \begin{cases}
\stenFwd{n}{d}{, \irrefl} & \;\;\textit{if} \;\, \neg \, r \\[-0.1em]
\stenFwd{n}{d}{}  & \;\;\textit{if} \;\,  r
\end{cases} \\
\textit{or} \;\;
& \stenFwdS{n}{d} := \stenFwd{n}{d}{}
\end{align*}
%
and similarly for \term{bwd} and \term{cen}.

Stencil specifications in \textit{spec} are considered
equal when they have the same modifiers and $\equiv$ equates
their regions, \ie{}:
\begin{align*}
(\textit{region} \equiv \textit{region'})
 \Rightarrow \quad\;\; & \texttt{stencil} \; [\textit{approx},] \; [\textit{mult},] \;
\textit{region} \\[-0.4em]
\quad \equiv \; & \texttt{stencil} \; [\textit{approx},] \;
            [\textit{mult},] \; \textit{region}'
\end{align*}
%

\paragraph{Overlapping regions with +}
The first nine
equations (with labels of the form $\ldots \texttt{+} \ldots$)
explain the notion of \emph{region overlapping} with sum
\term{+}. In the first six, $p$ and $q$ range
over booleans denoting the absence (true) or presence (false)
of \irrefl{}. These are combined via disjunction. For example,
the rule $(\textsc{B} \, \term{+} \,
\textsc{F})$ explains that the sum of \term{backward}
and \term{forward} regions in the same dimension and the same
depth is equivalent to a \term{centered} region of that depth.
If the \term{forward} part is non-pointed ($p =
\textit{false}$) but \term{backward} is pointed ($q =
\textit{true}$) then their sum is equal to a \term{centered}
region with no \irreflS{} modifier by the disjunction ($p \vee q =
\textit{true}$), \ie{}:
%%
\begin{align*}
& \stenCenS{n}{d}  \\[-0.2em]
\equiv \;\; & \stenFwdSR{n}{d}{\irrefl} \; \texttt{+} \; \stenBwdS{n}{d}
\end{align*}
%%
That is, the \term{fwd} region is non-pointed describing code which is missing
a zero-offset neighbourhood index (\eg{}, missing $\texttt{a(i)}$), but the addition of
the \term{bwd} has no non-pointed attribute and thus provides the
missing zero offset index.

The equation (\textsc{point}) provides an annihilation of
\irrefl{} using a further syntactic sugar.
We write $R^{\irreflS}$ to denote a region constant with an
non-pointed attribute, \eg{},
%%
\begin{align*}
\stenFwdS{n}{d}^{\irreflS}  := \,
\stenFwdSR{n}{d}{\irrefl}
\end{align*}
%%
The left-hand side of (\textsc{point}) has a sum of
products of two regions $R$ and $S$,
where in each component of the sum one of $R$
or $S$ has the non-pointed attribute. These cancel to give
just $R \, \term{*} \, S$.

\paragraph{Interaction between \term{+} and \term{*}}

The last seven equations (from (\textsc{\term{*}Dim}) inclusive), explain
the algebraic behaviour of \term{+} and \term{*}.  Together, we see
that \term{+} is an associative, commutative, idempotent binary
operator that distributes with \term{*}, which is associative and
commutative. This distribution of \term{*} over \term{+} is key as it
is used internally to give a normal form for stencil specifications
akin to Disjunctive Normal Form (DNF) (where \term{+} is
disjunction and \term{*} is conjunction). We revisit this
normalisation in \S~\ref{sec:inf-step4}.

The equation (\textsc{$\ast$Dim}) uses syntactic sugar
where $R^d$ and $S^d$ denote regions whose dimension is $d$.
This equation shows that a product of two regions on
the same dimension is equal to their addition.

@@

\subsection{Approximation}

Figure~\ref{fig:inequations} defines a notion of sub-specifications
via the relation $<:^r$ on \textit{region}
syntax and then $<:$ on \textit{spec} syntax inductively.
Both relations are reflexive and transitive (we omit
these equations) and congruent
with respect to \term{+}, \term{*}, \term{atLeast}, \term{atMost},
and \term{readOnce}. The \trule{\textsc{eq}} rule connects region
equations with inequations. The \trule{\textsc{over}} rule explains
the notion of spatial over-approximation via overlapping regions.
The \trule{\textsc{rep}} allows the modifier that
enforces that every index is read just once to be dropped
as an approximation; \trule{\textsc{bound}} gives
a connection between the \term{atLeast} and \term{atMost}
approximation modifiers.

\begin{theorem}[Soundness of approximation]
\begin{equation*}
\forall S, T, M, n . \quad
S <: T \; \wedge (\consA{M}{\interp{S}_n})
\Rightarrow (\consA{M}{\interp{T}_n})
\end{equation*}
\end{theorem}

\begin{figure}[t]
\begin{align*}
\hspace{-0.5em}
\begin{array}{c}
%%
%% INEQUATIONS ON REGION
%%
%\framebox{$\textit{region} <: \textit{region}'$} \\[0.85em]
%
%\dfrac{}{R <:^r R}\trule{\textsc{refl}} \qquad \dfrac{P <:^r R \quad R <:^r Q}{P <:^r
%  Q}\trule{\textsc{trans}}
%\qquad
%\\[1.5em]
\vspace{-0.4em}
\setlength{\arraycolsep}{0.1em}
\dfrac{\hspace{3em} m \leq n \qquad\qquad r \implies s \hspace{3em}}
{\begin{array}{rl}
\stenFwdSR{m}{d}{r} & <:^r \stenFwdSR{n}{d}{s}
%& \;\, ds \subseteq es \wedge n \leq m
\\
\wedge \; \stenBwdSR{m}{d}{r} & <:^r \stenBwdSR{n}{d}{s}
%& \;\, ds \subseteq es \wedge n \leq m
\\
\wedge \; \stenCenSR{m}{d}{r} & <:^r \stenCenSR{n}{d}{s}
%& \;\, ds \subseteq es \wedge n \leq m
\\
\wedge \; \stenBwdSR{m}{d}{r} & <:^r \stenCenSR{n}{d}{s} \\
\wedge \; \stenFwdSR{m}{d}{r} & <:^r \stenCenSR{n}{d}{s} \\[1.2em]
\end{array}}\trule{\textsc{over}} \\[-0.2em]
\hspace{-0.5em}
\dfrac{R_1 <:^r S_1 \quad R_2 <:^r S_2}
      {R_1 \, \texttt{+} \, R_2 <:^r S_1 \, \texttt{+} \, S_2}
\trule{\textsc{cong}\texttt{+}}
\dfrac{R_1 <:^r S_1 \quad R_2 <:^r S_2}
      {R_1 \, \texttt{*} \, R_2 <:^r S_1 \, \texttt{*} \, S_2}
\trule{\textsc{cong}\texttt{*}} \\[1.5em]
%%
%% INEQUATIONS ON SPEC
%%
%\framebox{$\textit{spec} <: \textit{spec}'$} \\[0.8em]
%
%
\dfrac{R <:^r S}{\term{atLeast}  S <: \term{atLeast}
  R}\trule{\texttt{L}}
\;\;
\dfrac{R <:^r S}{\term{atMost}  R <: \term{atMost}
  S}\trule{\texttt{M}} \\[1.5em]
\dfrac{R <:^r S}{R <: \term{atMost} \, S}\trule{\textsc{grow}} \quad
\dfrac{R <:^r S}{S <: \term{atLeast} \, R}\trule{\textsc{shrink}} \quad
 \\[1.5em]
%%
\dfrac{\mathsf{dims}(R) \cap \mathsf{dims}(S) = \emptyset}{R \ast S <: R}
\trule{\textsc{Gen$\ast$}}
\quad
\dfrac{R \equiv S}{R <: S}\trule{\textsc{eq}}
%%
\\[1.5em]
\dfrac{R <: S}{\term{readOnce} \, R <: \term{readOnce} \,
  S}\trule{\textsc{cong}\texttt{R}}
\;\;
\dfrac{}{\term{readOnce} \, R <: R}\trule{\textsc{rep}}
%\dfrac{}{R <: \term{atMost} R}
%\quad
%\dfrac{}{R <: \term{atLeast} R}\\[1.5em]
\end{array}
\end{align*}
\caption{Inequations $<:^r$ on \textit{regions} and
$<:$ on \textit{specs} (omitting reflexivity
and transitivity). }
\label{fig:inequations}
\vspace{-0.8em}
\end{figure}

>>>>

  Absolute index is a constant integer expression relative to the enclosing
  loop.
\end{definition}

@@

\begin{definition}[Stencil computations]
\label{def:stencil}
  Let $a$ be an array of dimensionality $n$ and $\overline{b}$ a
  collection of arrays of arbitrary (possible differing)
  dimensionalities.
  A \emph{stencil computation} comprises an iteration by a set $I$ of at
  most $n$ induction variables
  over a subset of the index space of $a$. Each step of the interation
  determs the elements of $a$ by an
  assignment $a(\bar{e}) = e_r$ where each $e \in
  \bar{e}$ is either a neighbourhood or constant index.
  For all arrays $b \in \bar{b}$ and array subscripts $b(\bar{e}')$ in
  $e_r$ and that flow to $e_r$, then each $e' \in \bar{e}'$ is also
  either a neighbourhood or constant index.

  We refer to the assignment $a(\bar{e}) = e_r$ and all the associated
  statements with dataflow to $e_r$ as the \emph{stencil kernel}.
\end{definition}
%
% specifying the shape of indexing expressions for
%arrays $\bar{v}$ that are simultaenously permitted and expected within
%the code to which the specification is given.
%
%\marginpar{\dnote{Is it worth having this informal consistency here?}}
\begin{definition}[Consistency, informal]
Specifications are associated to a stencil kernel
and are declared for one or more array variables
$\bar{v}$ within that kernel.
  A stencil specification defines a \emph{region} associated to an
  array variable $a$ against which a stencil kernel should be
  \emph{consistent}. Informally, the region defined by a specification
  is an $n$-dimensional rectilinear shape in a discrete space crossing the origin with respect to the induction
  variable. Consistency is then a two way requirement.
  On the one hand, all array subscripts on $a$ contributing to the kernel
  should be consistent with the region; on the other, all parts of the
  region should have an index corresponding to that part.
\end{definition}

@@

Figure~\ref{fig:syntax} gives the syntax of stencil specifications,
which we introduce in stages below.  The top-level is given by the
\nonterm{specification} production which splits into either a
\nonterm{regionDec} (region declaration) or a \nonterm{specDec}
(specification declaration).

@@

Specifications have a theory of equality $\equiv$ as well
as approximation $<:$ which we define here.
For brevity we use abbreviations $\term{fwd}$, $\term{bwd}$, and
$\term{cen}$ for region constants here.

The equality relation $\equiv$ is defined over \emph{region}
syntax which gives the following axioms on regions:
%
\begin{itemize}
\item $\term{+}$ is idempotent, commutative, and associative;
\item $\term{*}$ is commutative, associative, and distributes over
$\term{+}$, \ie{},  $R \; \texttt{*} \; (S \; \texttt{+} \; T) \; \equiv \; (R \;
                       \texttt{*} \; S) \; \texttt{+} \; (R
                       \; \texttt{*} \; T)$. 
Distributivity is 
used later to give a normal form for specifications
akin to Disjunctive Normal Form (\S~\ref{sec:inf-step4}).

\item (\emph{overlapping}) Given two region constants $R_1$ and $R_2$ 
on the same dimension, then some combinations of region constants
can be coalesced into a single region constant $R_3$ if $R_1$ completely
overlaps $R_2$. There are five possible instantiations of the
following rule $R_1 + R_2 \equiv R_3$ which we denote: $\term{fwd + fwd} \equiv \term{fwd}$, $\term{bwd + bwd} \equiv \term{bwd}$, 
$\term{cen + cen} \equiv \term{cen}$, $\term{cen + fwd} \equiv \term{cen}$ and
$\term{cen + bwd} \equiv \term{cen}$
%
\newcommand{\stenGR}[4]{\small{\textit{#4}(\term{depth=}#1,
  \term{dim=}#2, #3)}}
\newcommand{\stenG}[3]{\small{\textit{#3}(\term{depth=}#1,
  \term{dim=}#2)}}
\begin{align*}
\hspace{-1em} & \stenGR{n}{d}{p \vee q}{$R_3$} \\[-0.4em]
\hspace{-1em} \equiv \; & \stenGR{n}{d}{p}{$R_1$} \; \texttt{+} \; \stenGR{m}{d}{q}{$R_2$}
\end{align*}
where $n \geq m$ and $p$ and $q$ range
over booleans representing the absence (true) or presence (false)
of \irrefl{} which are combined via disjunction.

\item (\textit{overlapping pointed}) \term{pointed} regions coalesce with other regions of the
same dimension:
%
\begin{align*}
& \stenG{n}{d}{region} \\[-0.4em]
\equiv \; & \stenGR{n}{d}{p}{region} \; \texttt{+} \;
                          \stenReflS{d}
\end{align*}
%
\item \textit{(centered)} a forward and backward region of the same
  depth and dimension coalseces into a centered region:
\begin{align*}
\hspace{-1em} & {\small{\stenCen{n}{d}{, p \vee q}}} \\[-0.4em]
\hspace{-1em} \equiv \; & {\small{\stenFwdSR{n}{d}{p} \; \texttt{+} \;
                          \stenBwdSR{n}{d}{q}}}
\end{align*}
%%
\item $(\textsc{point}) \;\; (R \, \term{*} \, S^\irreflS) \, \term{+} \,
                         (R^\irreflS \, \term{*} \, S) \; \equiv \; R
                         \, \term{*} \, S$
meaning that for the sum of products of two regions $R$ and $S$,
where in each component one of $R$ or $S$ has the non-pointed
attribute denoted by $R^{\term{np}}$ and $S^{\term{op}}$,
the non-pointed attributes cancel to give $R \, \term{*} \, S$.
%
\item $(\textsc{$\ast$Dim}) \;\; R^d \, \term{$\ast$} \, S^d \equiv R^d
                             \, \term{+} \, S^d$
%
where $R^d$ and $S^d$ denote regions whose dimension is $d$.
Thus a product of two regions on
the same dimension is equal to their addition.
\end{itemize}
%
Stencil specifications in \textit{spec} are considered
equal when they have the same modifiers and $\equiv$ equates
their regions.

@@

Figure~\ref{fig:inequations} defines a notion of sub-specifications
via the relation $<:^r$ on \textit{region}
syntax and then $<:$ on \textit{spec} syntax inductively.
Both relations are reflexive and transitive (we omit
these equations). The $<:^r$ relation is congruent
with respect to \term{+} and \term{*} $<:$ is congruent
with repect to \term{readOnce}, the equations of which we omit for brevity.

The \trule{\textsc{eq}} rule connects region
equations to approximation; \trule{\textsc{over}} explains
the notion of spatial over-approximation via overlapping regions; 
\trule{\textsc{rep}} allows the \term{readOnly} modifier that
to be dropped. The \trule{\textsc{shrink}} 
and \trule{\textsc{grow}} rules turn the spatial approxiamation on
regions into approximation on specifications via the \term{atLeast} and \term{atMost}
modifiers. Finally, \trule{\textsc{Gen$\ast$}} shows
that specifications can be generalised by removing parts of a product
$\ast$ which have disjoint dimensions to the rest of the region
product (where $\mathsf{dims}$ denotes the set of dimensions specified
in a region).

\begin{theorem}[Soundness of approximation]
A piece of code consistent with a specification $S$
 is consistent with a generalisation $T$ of the specification.
That is for all models $M$ of a piece of code:
\begin{equation*}
S <: T \; \wedge (\consA{M}{\interp{S}_n}) \;
\Rightarrow \; (\consA{M}{\interp{T}_n})
\end{equation*}
where $\consAName$ is the consistency predicate used in checking (\S~\ref{subsec:checking}).
\end{theorem}

\begin{figure}[t]
\vspace{-0.5em}
\begin{align*}
\hspace{-0.5em}
{\small{
\begin{array}{c}
%%
%% INEQUATIONS ON REGION
%%
%\framebox{$\textit{region} <: \textit{region}'$} \\[0.85em]
%
%\dfrac{}{R <:^r R}\trule{\textsc{refl}} \qquad \dfrac{P <:^r R \quad R <:^r Q}{P <:^r
%  Q}\trule{\textsc{trans}}
%\qquad
%\\[1.5em]
\vspace{-0.4em}
\setlength{\arraycolsep}{0.1em}
\dfrac{\hspace{3em} m \leq n \qquad\qquad r \implies s \hspace{3em}}
{\begin{array}{rl}
\stenFwdSR{m}{d}{r} & <:^r \stenFwdSR{n}{d}{s}
%& \;\, ds \subseteq es \wedge n \leq m
\\
\wedge \; \stenBwdSR{m}{d}{r} & <:^r \stenBwdSR{n}{d}{s}
%& \;\, ds \subseteq es \wedge n \leq m
\\
\wedge \; \stenCenSR{m}{d}{r} & <:^r \stenCenSR{n}{d}{s}
%& \;\, ds \subseteq es \wedge n \leq m
\\
\wedge \; \stenBwdSR{m}{d}{r} & <:^r \stenCenSR{n}{d}{s} \\
\wedge \; \stenFwdSR{m}{d}{r} & <:^r \stenCenSR{n}{d}{s} \\[1.2em]
\end{array}}\trule{\textsc{over}} \\[-0.2em]
%\hspace{-0.5em}
%\dfrac{R_1 <:^r S_1 \quad R_2 <:^r S_2}
%      {R_1 \, \texttt{+} \, R_2 <:^r S_1 \, \texttt{+} \, S_2}
%\trule{\textsc{cong}\texttt{+}}
%\dfrac{R_1 <:^r S_1 \quad R_2 <:^r S_2}
%      {R_1 \, \texttt{*} \, R_2 <:^r S_1 \, \texttt{*} \, S_2}
%\trule{\textsc{cong}\texttt{*}} \\[1.5em]
%%
%% INEQUATIONS ON SPEC
%%
%\framebox{$\textit{spec} <: \textit{spec}'$} \\[0.8em]
%
%
\dfrac{R <:^r S}{\term{atLeast}  S <: \term{atLeast}
  R}\trule{\texttt{L}}
\;\;
\dfrac{R <:^r S}{\term{atMost}  R <: \term{atMost}
  S}\trule{\texttt{M}} \\[1.5em]
\dfrac{R <:^r S}{R <: \term{atMost} \, S}\trule{\textsc{grow}} \,
\dfrac{R <:^r S}{S <: \term{atLeast} \, R}\trule{\textsc{shrink}} \,
\dfrac{R \equiv S}{R <:^r S}\trule{\textsc{eq}}
 \\[1.5em]
%%
\dfrac{\mathsf{dims}(R) \cap \mathsf{dims}(S) = \emptyset}{R \ast S <: R}
\trule{\textsc{Gen$\ast$}}
\quad
%%
%\\[1.5em]
%\dfrac{R <: S}{\term{readOnce} \, R <: \term{readOnce} \,
%  S}\trule{\textsc{cong}\texttt{R}}
%\;\;
\;
\dfrac{}{\term{readOnce} \, R <: R}\trule{\textsc{rep}}
%\dfrac{}{R <: \term{atMost} R}
%\quad
%\dfrac{}{R <: \term{atLeast} R}\\[1.5em]
\end{array}}}
\end{align*}
\caption{Inequations $<:^r$ on \textit{regions} and
$<:$ on \textit{specs} (omitting reflexivity, transitivity, and congruences). }
\label{fig:inequations}
\vspace{-0.8em}
\end{figure}
